Proportional multi-state multiple-cohort life table model

### 2018-05-08 MSLT/MSLTdoc.Rmd
Checked total Lx and LXw calcs

## Uncommnet to check life table list
# View(general_life_table_list_bl[[2]])

@@

i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
i_sex <- c("male", "female")
i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)

@@

## Uncommnet to check scenario incidence
# View(incidence_sc[[1]])
```

@@

## Generate total change in mortality rate
## Sum mortality rate scenarios (mx_sc_total)

@@

        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
        index <- index + 1
      }
    }
    mx_sc_total[[l_index]] <- mortality_sum 
    l_index <- l_index + 1
  }
}

## Uncommnet to check sceanrio mortality and changes 
# View(mx_sc_total[[1]])

@@

for (age in i_age_cohort){
  for (sex in i_sex){
    # cat("age ", age, " and sex ", sex, "\n")
    # modify idata's mortality and pyld total for the said scenario
    td2 <- input_life_table
    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx 
                                                            + mx_sc_total[[index]]$total
    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate 
                                                                    + pylds_sc_total[[index]]$total

    # Instead of idata, feed td to run scenarios
    general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)

    index <- index + 1
  }
}
## Uncommnet to check scenario life tables
# View(general_life_table_list_sc[[1]])
# View(general_life_table_list_bl[[2]])

@@

## Uncomment to check
# View(output_burden[[32]])

@@

                jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
                                          mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

@@

               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

@@

graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
          knitr::include_graphics(graphs_doc [[index]])
          index <- index + 1
        }
    }
}
 knitr::include_graphics(graphs_doc[c(1:64)])
```

@@

```{r, echo=TRUE}
####Generate data frame with all outputs to graph total change in burden by simulation year. 
###first, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to 
####include. To show all select 16. what is specified in in_outcomes will be graphed or can also be presented as a total. 
aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))

@@

#####The following adds up both males and females
# Remove non-numeric columns starting with age and sex
aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))

@@

# Create a copy of aggregate_frame_females
total_aggr <- aggregate_frame_females
# Add aggregate_frame_males values to it
for (i in 1:ncol(aggregate_frame_females)){
  total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
}
total_aggr$sim_year <- seq.int(nrow(total_aggr))
```

25) Plots using total aggregated change in burden.

@@

disease_outcome <- "Health adjusted life years"

@@

  geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
  theme_classic() +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
  geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
  theme(plot.title = element_text(hjust = 0.5))

>>>>

## Uncommnet to check life table list
# View(general_life_table_list_bl[[32]])

@@

i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")

@@

## Uncommnet to check scenario incidence
# View(incidence_sc[[1]])

@@

## Generate total change in mortality rate

## Sum mortality rate change scenarios (mx_sc_total)

@@

        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
        index <- index + 1
      }
    }
    mx_sc_total[[l_index]] <- mortality_sum

    l_index <- l_index + 1
  }
}

## Uncommnet to check sceanrio mortality and changes 
# View(mx_sc_total[[3]])

@@

for (age in i_age_cohort){
  for (sex in i_sex){

    cat("age ", age, " and sex ", sex, "\n")
    # modify idata's mortality and pyld total for the said scenario
    td2 <- input_life_table
    # td2 <- subset(td2, select = -c(mx, pyld_rate))
    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total

    # Instead of idata, feed td to run scenarios
    general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)
    #

    index <- index + 1
  }
}
## Uncommnet to check scenario life tables
# View(general_life_table_list_sc[[32]])
# View(general_life_table_list_bl[[1]])

## Check difference life table baseline and scenario
general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx

@@

## Uncomment to check
# View(output_burden[[1]])

@@

  tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
  grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
                                          mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

@@

               tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

@@

graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".tiff", sep = ""))
          knitr::include_graphics(graphs_doc [[index]]) 
          # %>%
          #   set_names(paste("Change in", ifelse(outcome == "mx", "Deaths", "Incidence"), "over time", sep = ""))

          index <- index + 1
        }
    }
}
```

```{r, echo=FALSE}
## fig.cap=paste(names(graphs_doc)), message=FALSE, 
 knitr::include_graphics(graphs_doc[c(1:64)])
```

@@

```{r, echo=TRUE}
## Generate data frame with all outputs to graph total change in burden by simulation year. 
## First, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to include. To show all select 16. what is specified in in_outcomes will be graphed or can also be presented as a total.

## DO A LIST TO GENERATE ALL OUTCOMES ALL. CHECK SUMMARION OF COHORTS ORDER.

aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))

@@

## The following adds up both males and females
## Remove non-numeric columns starting with age and sex
aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))

@@

## Create a copy of aggregate_frame_females
total_aggr <- aggregate_frame_females
## Add aggregate_frame_males values to it
for (i in 1:ncol(aggregate_frame_females)){
  total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
}
total_aggr$sim_year <- seq.int(nrow(total_aggr))

## Uncomment to view aggregated data
sum(total_aggr$total_Lx_diff)
```

25) Plots using total aggregated change in burden. This graphs are not very good for life years and health adjusted life years as the difference between baseline and scenario is very small relative to the total, hence the graphs show almost no change.

@@

disease_outcome <- "Life years"

@@

  geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
  theme_classic() +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
  geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(breaks=seq(0,7000000,by=1000000))
